### 2023-04-21 SFE-article/Sections/Parameter_estimation.tex
21.04

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

The goal of parameter estimation is to obtain the "best" estimate of parameter set $\theta$ (which is a subset of the parameter space $\Theta$ containing all parameters of a model) based on the continuous observations $Y(t)$ or the discrete observations $Y(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, $y(t)$ (equation \ref{Model_measurment_1}), to give the observable dependent variable $Y(t)$ (for example results of an experiments). For discrete observations, this can be expressed as:

{\footnotesize
	\begin{equation*}
		Y(t_i) = y(\theta, t_i) + \epsilon(t_i)
\end{equation*} }

For continuous variables, the equation is:

{\footnotesize
	\begin{equation*}
		Y(t) = y(\theta, t) + \epsilon(t)
\end{equation*} }

However, obtaining analytical solutions for a deterministic process model can be challenging, so experiments are often conducted where the vector of derivatives $dY(t_i)/dt$ is measured instead of $Y(t_i)$ itself. In such cases, it is assumed that the unobservable error is added to the deterministic derivative $dy(\theta, t_i)/dt$ (as shown in equation \ref{Model_measurment_2}) as follows:

{\footnotesize
	\begin{equation*}
		\cfrac{d Y(t_i)}{dt} = \cfrac{dy(\theta, t_i)}{dt} + \epsilon(t_i)
\end{equation*} }

In the case where the error in the first observation is denoted as $\epsilon_1$, the error in the second observation $\epsilon_2'$ incorporates $\epsilon_1$ as well as an independent random component, given by $\epsilon_2' = \epsilon_1 + \epsilon_2$. Similarly, the error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so on. \citet{Mandel1957}  made a distinction between the typically assumed independent measurement error in the dependent variable and a "cumulative" or interval error, in which each new observation encompasses the error of the previous ones. Cumulative errors arise from fluctuations in the process itself due to small variations in operating conditions and are not independent; only the differences in measurement from one period to the next are independent.

\subsubsection{Least squares estimation}
In statistics, least squares parameter estimation is a method used to estimate the parameters of a linear regression model. One of the benefits of this method is that it does not require any prior knowledge about the distribution of unobservable errors. Additionally, it yields unbiased estimates and results in the minimum variance among all linear unbiased estimators.

If the model responses, represented by $Y$, are continuous functions of time from $t = 0$ to $t = t_f$, the Markov criterion, also known as "rigorous least squares," is used to minimize the integral squared error $\gamma$. This is achieved by minimizing the following expression:

{\footnotesize
	\begin{align}
		\gamma &= \cfrac{1}{2} \int_{0}^{t_f} \left[ Y(t) - y(\theta,t) \right]^T \Gamma^{-1} \left[ Y(t) - y(\theta,t) \right] dt \\
		\hat{\theta} &= \arg \min_{\theta \in \Theta} \gamma
\end{align} }

Here, $\Gamma$ represents the covariance matrix or a matrix of appropriate weights, $\hat{\theta}$ is the parameter estimate, $\Theta$ is the parameter space, and $\gamma$ is the time-integrated value of the error squared.

Alternatively, if the observations are made at discrete time intervals $t_i$, where $i = 1, 2, \dots, n$, the Markov criterion is to minimize the sum of squared errors, which is given by:

{\footnotesize
	\begin{align}
		\gamma &= \cfrac{1}{2} \sum_{i=1}^{n} \left[ Y(t_i) - y(\theta,t_i) \right]^T \Gamma^{-1} \left[ Y(t_i) - y(\theta,t_i) \right] \\
		\hat{\theta} &= \arg \min_{\theta \in \Theta} \gamma
\end{align} }

The parameter estimate, $\hat{\theta}$, is obtained by minimizing $\gamma$ over the parameter space $\Theta$.

If the covariance matrix, $\Gamma$, is a diagonal matrix, the Markov criterion becomes a "weighted least squares" criterion. If $\Gamma = \sigma^2 \textbf{I}$, where $\sigma$ is a standard deviation and $\textbf{I}$ is the identity matrix, the Markov criterion becomes the "ordinary least squares" criterion.

\iffalse
\subsubsection{Weighted sum of squares - Cumulative data}

One characteristic feature of certain special experiments is the use of the same batch of materials for the entire series of measurements. If unobservable error in the first observation is designated as $\epsilon_1$, the second error observation $\epsilon_2'$ includes $\epsilon_1$ plus a random component introduced side from $\epsilon_1$, or $\epsilon_2' = \epsilon_1 + \epsilon_2$. The error in the third observation is $\epsilon_3' = \epsilon_1 + \epsilon_2 + \epsilon_3$, and so forth. \citet{Mandel1957} distinguished between the usually assumed type of independent measuring error in the dependent variable and a 'cumulative' or interval error in which each new observation includes the error of the previous observations. Cumulative errors, arising because of the fluctuations as a function of time in the process itself due to small changes in operating, are not independent - only the differences in measurement from one period to the next are independent. Thus, if we consider two model:

{\footnotesize
	\begin{flalign*}
		&{\normalsize \text{Model A - independent error:} } &&\nonumber \\
		Y(t_i) &= y(\theta, t_i) + \epsilon_i  \qquad i = 1,2,...,n&& \\
		&{\normalsize \text{Model B - cumulative data:}} && \nonumber \\
		Y(t_j) - Y(t_{j-1}) &= y(\theta, t_j) - y(\theta, t_{j-1}) + \epsilon_j, \quad j = 1,2,...,n&& \\
		&{\normalsize \text{or}} && \nonumber \\
		Y(t_i) = \sum_{j=1}^{i} &\left( Y(t_j) - Y(t_{j-1}) \right) = \sum_{j=1}^{i} \left( y(\theta, t_j) - y(\theta, t_{j-1}) \right) + \sum_{j=1}^{i} \epsilon_j &&
\end{flalign*} }

in which $\epsilon_i$ are independent random variables with $\mathbb{E}\{\epsilon_i\}$ and Var$\{\epsilon_i\} = \sigma_1^2$, a constant. The $\epsilon_j$ are also random independent variables because they represent differences such as $\epsilon_3' - \epsilon_2' = \left( \epsilon_1 + \epsilon_2 + \epsilon_3 \right) - \left( \epsilon_1 + \epsilon_2 \right) = \epsilon_3$. We assume that $\mathbb{E}\{\epsilon_j\}$ and Var$\{\epsilon_i\} = \sigma_1^2\left(y_j - y_{j-1}\right)$; i.e., the variance of $\epsilon_j$ can be a function of the test interval. The experimental point tend to stay on one side of the regression line of the best fit for Model B.

The unbiased estimate of the Model B is obtained by minimizing the weighted sum of squares:

{\footnotesize
	\begin{equation*}
		\hat{\theta} = \arg \min_{\theta \in \Theta} \sum_{j=1}^{n} \cfrac{ \left[ \left( Y(t_j) - Y(t_{j-1}) \right) - \left( y(\theta, t_j) - y(\theta, t_{j-1}) \right) \right]^2  }{y(\theta, t_j) - y(\theta, t_{j-1})}
\end{equation*} }

\fi

\subsubsection{Maximum likelihood estimation}

Maximum likelihood estimation (MLE) is a method of estimating the parameters of an assumed probability distribution, given some observed data. This is achieved by maximizing a likelihood function so that, under the assumed statistical model, the observed data is most probable. The MLE has the desirable characteristics of asymptotic efficiency and normality. Each time it has been associated with the (joint) normal distribution because of mathematical convenience. Consider the joint probability density function (the likelihood function) $p \left(\theta | y(t_1),y(t_2),...,y(t_n)\right)$ for parameter $\theta$. If a maximum of this function over all choices of $\theta$ can be found, the estimates so obtained are maximum likelihood estimates. The conditions at the maximum can be evolved incorporating prior information as follows.

The posterior probability density $p(\theta|y)$ can be expressed as the ratio of two probability densities if we make use of the analogue for continuous variables of Equation \ref{EQ: Probabiltiy_independet_b}:

{\footnotesize
	\begin{align} \label{EQ: Posterior_distribution}
		p\left( \theta| y(t_n),...,y(t_1) \right) = \cfrac{p\left( \theta, y(t_n),...,y(t_1) \right)}{p\left( y(t_n),...,y(t_1) \right)}
\end{align} }

\todo[]{Notation is a bit unclear}The numerator of the right-hand side od Equation \ref{EQ: Posterior_distribution} using Equation \ref{EQ: Probabiltiy_dependet_a} becomes

{\footnotesize
	\begin{equation}
		\begin{split}
			p\left( \theta,y(t_n),...,y(t_1) \right) &= p\left( y(t_n)|\theta,y(t_{n-1}),...,y(t_1) \right) \\ &\cdot p\left( \theta,y(t_{n-1}),...,y(t_1) \right)
		\end{split}
\end{equation} }

These operations can be continued repetitively until we get

{\footnotesize
	\begin{equation}
		p\left( \theta, y(t_n),...,y(t_1) \right) = p\left( \theta \right) \prod_{i=1}^{n} p\left( y(t_i)|\theta, y(t_{i-1}),...,y(t_1) \right)
\end{equation} }

Examination of Equation \ref{EQ: Measurment_noise} shows that $ Y(t_i)$ depends only on $t_i$, $\theta$ and $\epsilon(t_i)$ and is not conditioned by any previous measurement. Consequently, we can write

{\footnotesize
	\begin{equation}
		p\left( y(t_i)|\theta, y(t_{i-1}),...,y(t_1) \right) = p\left( y(t_i)|\theta \right)
\end{equation} }

provided Equation \ref{EQ: Measurment_noise} is observed as a constraint. The desired joint conditional probability function is thus

{\footnotesize
	\begin{equation}
		p\left( \theta| y(t_n),...,y(t_1) \right) = \cfrac{p\left( \theta \right) \prod_{i=1}^{n} p\left( y(t_i)|\theta \right) }{p\left( y(t_n),...,y(t_1) \right)}
\end{equation} }

We can get rid of the evidence term $p\left( y(t_n),...,y(t_1) \right)$ because it's constant with respect to the maximization. Moreover, if we are lacking a prior distribution over the quantity we want to estimate, then $p(\theta)$ can be omitted. In such a case:

{\footnotesize
	\begin{equation}
		p\left( \theta| y(t_n),...,y(t_1) \right) = \prod_{i=1}^{n} p\left( y(t_i)|\theta \right) = \prod_{i=1}^{n} L\left( \theta|y(t_i) \right) 
\end{equation} }

By collecting a values of $y$ and selecting the values of $\theta$ that maximize the likelihood function $L\left( \theta | y \right)$, a function described in chapter \ref{CH: Bayes} in connection with Bayes' theorem\todo{Is the reference it previous chapter needed?}. Such estimators, $\hat{\theta}$, are known as maximum likelihood estimators. In effect, the methods selects those values of $\theta$ that are at least as likely to generate the observed sample as any other set of values of the parameters if the probability density of $y$ were to be extensively simulated through use of the probability density $p\left( y|\theta \right)$. In making a maximum likelihood estimate, we assume the form of the probability density and that all possible values of $\theta$ are equally likely before experimentation. \todo{Needed?}

%The likelihood function for the parameters given one observation is just the probability density in which the observation is regarded as a fixed number and the parameters as the variables:
%
%{\footnotesize
	%	\begin{equation}
		%		p\left( \theta| y(t_i) \right) = L\left( \theta|y_{t_i} \right) = p\left( y(t_i)| \theta \right)
		%\end{equation} }
		%where the lower case $y$'s and the number subscripts designate the value of the respective observation that is inserted into the probability density function. 
		The likelihood function for the parameters based on several observations is the product of the individual functions if the observations are independent.

		{\footnotesize 
			\begin{multline}
				L\left( \theta|y(t_n),...,y(t_1) \right) = \prod_{i=1}^{n} L\left( \theta|y(t_i) \right) \\
				= p\left( y(t_1)| \theta \right)p\left( y(t_2)| \theta \right) ... p\left( y(t_n)| \theta \right)
		\end{multline} }

		In choosing as estimates of $\theta$ the values that maximize $L$ for the given values $\left( y(t_i) \right)$, it turns out that it is more convenient to work with the $\ln L$ than with $L$ itself:

		{\footnotesize 
			\begin{equation}
				\ln L = \ln p \left( y(t_1)| \theta \right) + \ln p\left( y(t_2)| \theta \right) + ... + \ln p \left( y(t_n)| \theta \right) = \sum_{i=1}^{n} \ln p\left( y(t_i); \theta \right)
		\end{equation} }

		By assuming that the conditional distribution of $\bar{Y}_i$, given $y_i$, is normal, then we form the likelihood function based on the probability density:

		{\footnotesize
			\begin{align} \label{EQ: MLE_Norm}
				p\left(  \theta, \sigma | y(t_n),...,y(t_1) \right) &= \prod_{i=1}^{n} \cfrac{1}{\sqrt{2 \pi} \sigma } \exp \left[ \cfrac{1}{2  \sigma } \left( Y(t_i) - y(\theta , t_i) \right)^2 \right] \nonumber \\
				L\left( \theta, \sigma | y(t_n),...,y(t_1) \right) &= \prod_{i=1}^{n} \cfrac{1}{\sqrt{2 \pi} \sigma } \exp \left[ \cfrac{1}{2  \sigma } \left( Y(t_i) - y(\theta , t_i) \right)^2 \right]
		\end{align}}
		where $\sigma$ is the variance

		%In Equation \ref{EQ: MLE_Norm} the parameters are the variables, and the values of $Y$ and $x$ are given. Then
		By taking the natural logarithm of the Equation \ref{EQ: MLE_Norm}, the final form of the objective function can be obtained:

		{\footnotesize
			\begin{equation}
				\ln L = -\cfrac{n}{2} \left( \ln \sqrt{2 \pi} + \ln \sigma^2 \right)
				- \cfrac{ \sum_{i=1}^{n} \left[  Y(t_i) - y(\theta , t_i) \right]^2 }{2 \sigma^2}
			\end{equation}
		}

		The parameter estimation problem can be formulated as follow:

		{\footnotesize
			\begin{equation}
				\begin{aligned} \label{EQ: Optimization_formulation_MLE}
					&\hat{\theta}_{MLE} &= \arg \max_{\sigma, \theta \in \Theta} \ln L = \arg \max_{\sigma,\theta \in \Theta} p(\theta|y) \\
					&\text{subject to}
					& \dot{x} = f(t,x,\theta) \\
					&& \dot{\theta} = 0 \\
					&& y = y(x)
				\end{aligned}
		\end{equation} }

		\todo{$x$ has not been defined}

		Based on the first order optimality condition, the $\ln L$ can be maximized with respect to the vector $\theta$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

		{\footnotesize
			\begin{align}\label{EQ: MLE_Basic_Solution}
				\cfrac{\partial \ln L}{\partial \theta} &= \cfrac{\partial \sum_{i=1}^{n} \ln p\left( y(t_i) | \theta \right) }{\partial \theta} = 0 
		\end{align} }

		Solution of Equations \ref{EQ: MLE_Basic_Solution} yield the desired estimates $\hat{\theta}$. For some models, these equations can be explicitly solved for $\hat{\theta}$ but in general no closed-form solution to the maximization problem is known or available, and an MLE can only be found via numerical optimization.

		\iffalse

		It can be shown that as $n$ approaches infinity the maximum likelihood estimates have the desirable asymptotic properties:

		\begin{enumerate}[label=(\arabic*)]
			\item {\footnotesize $\lim\limits_{n \rightarrow \infty} \mathbb{E}\{\hat{\theta}\} = \theta$ }
			\item {\footnotesize $\left[ \sqrt{n}\left( \hat{\theta} - \theta \right) \right]$ } is normally distributed
		\end{enumerate}
		and for the case of two parameters:

		\begin{enumerate}[resume,label=(\arabic*)]
			\item {\footnotesize $\lim\limits_{n \rightarrow \infty} \text{Var}\{\hat{\theta}\} = \cfrac{1}{n} \left[ \mathbb{E} \left\{ \left( \cfrac{\partial \ln p}{\partial \theta} \right)^2 \right\} \right]^{-1} \cfrac{1}{1-\rho^2_{\hat{\theta}}} $ } \label{Assumption_MLE}
		\end{enumerate}

		where $\rho^2_{\hat{\theta}}$ is the matrix of correlation coefficients.

		Maximum likelihood estimates are not necessarily unbiased. Maximum likelihood estimates, however are efficient and, hence, consistent estimates. Furthermore, where a sufficient estimate can be obtained, the maximum likelihood method will obtain it. Finally, if $\hat{\theta}$ is a maximum likelihood estimator of $\theta$, the $f\left( \hat{\theta} \right)$ is a maximum likelihood estimator of $f\left( \theta \right)$, a function of $\theta$.

		\hrule

		%Under the assumption of normality, the lack of prior knowledge about the parameters $\theta^0$ and known initial conditions $y_0$. Intuitively, this selects the parameter values that make the observed data most probable. The specific value $\hat{\theta}$ that maximizes the likelihood function $L$ is called the maximum likelihood estimate.

		As already discussed, the prior information about the estimated parameters can be incorporated in to the maximum likelihood estimator to form a maximum a posteriori probability (MAP) estimate. MAP can be used to obtain a point estimate of an unobserved quantity on the basis of empirical data. MAP estimation can therefore be seen as a regularization of maximum likelihood estimation. The relationship between MAP and MLE can be obtained from the Bayes' rule:

		{\footnotesize
			\begin{equation} 
				\begin{split} \label{EQ: MAP_probability}
					{\bf \hat{\theta}_{\text{MAP}}} &= \arg\max_{\sigma, \theta \in \Theta} p(\theta | y) \\
					&= \arg\max_{\sigma, \theta \in \Theta} \frac{p(y | \theta) p(\theta)}{p(y)} \\
					&= \arg\max_{\sigma, \theta \in \Theta} p(y | \theta) p(\theta) \\
					&= \arg\max_{\sigma, \theta \in \Theta} \ln(p(y | \theta) p(\theta)) \\
					&= \arg\max_{\sigma, \theta \in \Theta} \ln p(y | \theta) + \ln p(\theta)
				\end{split}
		\end{equation} }

		We assume the normal distribution of random samples, with identical variance and the prior distribution of $\theta$ given by $\mathcal{N}(\theta^0,\,\sigma_\theta^{2})$. The function to be maximized can be written as

		{\footnotesize
			\begin{equation} \label{EQ: MAP_L2}
				\begin{aligned}
					&\arg\max_{\sigma, \theta \in \Theta} \left[ \ln \prod_{i=1}^{n} \frac{1}{\sigma\sqrt{2\pi}}e^{-\frac{ \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2 }{2\sigma^2}}
					+ \ln \prod_{j=0}^{p} \frac{1}{\sigma_\theta\sqrt{2\pi}}e^{-\frac{ \left( \theta - \theta^0 \right)^2}{2\sigma_\theta^2}} \right]  \\
					= &\arg\max_{\sigma, \theta \in \Theta} \left[- \sum_{i=1}^{n} {\frac{\left( Y(t_i) - y\left( \theta, t_i \right) \right)^2}{2\sigma^2}}
					- \sum_{j=0}^{p} {\frac{\left( \theta - \theta^0 \right)^2}{2\sigma_\theta^2}} \right]\\
					= &\arg\min_{\sigma, \theta \in \Theta} \frac{1}{2\sigma^2} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2
					+ \frac{\sigma^2}{\sigma_\theta^2} \sum_{j=0}^{p} \left( \theta - \theta^0 \right)^2 \right] \\
					= &\arg\min_{\sigma, \theta \in \Theta} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2 + \lambda_{L2} \sum_{j=0}^{p} \left( \hat{\theta} - \theta^0 \right)^2 \right]
				\end{aligned}
		\end{equation} }

		In Equation \ref{EQ: MAP_L2} some of the constants (with respect to $\theta$) were dropped and factored to simplify the expression. We can adjust the amount of regularization we want by changing $\lambda_{L2}$. Equivalently, we can adjust how much we want to weight the priors carry on the coefficients ($\theta$). If we have a very small variance (large $\lambda_{L2}$) then the coefficients will be very close to 0; if we have a large variance (small $\lambda_{L2}$) then the coefficients will not be affected much (similar to as if we didn't have any regularization).

		Let's assume that the prior distribution of parameter $\theta$ is given by the Laplace distribution. The likelihood function can defined as follow:

		{\footnotesize
			\begin{equation} \label{EQ: MAP_L1}
				\begin{aligned}
					&\arg\max_{\sigma, \theta \in \Theta} \left[ \ln \prod_{i=1}^{n} \frac{1}{\sigma\sqrt{2\pi}}e^{-\frac{ \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2 }{2\sigma^2}}
					+ \ln \prod_{j=0}^{p} \frac{1}{2b}e^{-\frac{ | \theta - \theta^0 | }{\sigma_\theta}} \right]  \\
					= &\arg\max_{\sigma, \theta \in \Theta} \left[- \sum_{i=1}^{n} {\frac{\left( Y(t_i) - y\left( \theta, t_i \right) \right)^2}{2\sigma^2}}
					- \sum_{j=0}^{p} {\frac{ | \theta - \theta^0 | }{\sigma_\theta}} \right]\\
					= &\arg\min_{\sigma, \theta \in \Theta} \frac{1}{2\sigma^2} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2
					+ \frac{\sigma^2}{\sigma_\theta} \sum_{j=0}^{p} | \theta - \theta^0 | \right] \\
					= &\arg\min_{\sigma, \theta \in \Theta} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \theta, t_i \right) \right)^2 + \lambda_{L1} \sum_{j=0}^{p} | \theta - \theta^0 | \right]
				\end{aligned}
		\end{equation} }

		The Laplacean prior act as $L1$-type of regularization on the optimization problem. Instead of preventing any of the coefficients from being too large (due to the squaring), $L1$ promotes sparsity. Due to the sparsity some of the parameters can be zero out, and eventually removed from the model. $L1$ is more robust to outliers than $L2$.

		If the cost function is given by Equations \ref{EQ: MAP_L2} or \ref{EQ: MAP_L1}, then the optimization problem can be formulated as follow:

		{\footnotesize
			\begin{equation}
				\begin{aligned}
					&\hat{\theta}_{MAP} &= \arg\max_{\sigma, \theta \in \Theta} \ln p(y | \theta) + \ln p(\theta) \\
					&\text{subject to}
					& \dot{x} = f(t,x,\theta) \\
					&& \dot{\theta} = 0 \\
					&& y = g(x)
				\end{aligned}
		\end{equation} }

		\hrule
		If the cost function to be maximized is given be Equation \ref{EQ: MAP_probability} and constrained by Equation \ref{EQ: Measurment_noise}, then the method of Lagrange multipliers can be used to convert a constrained problem into a form such that the derivative test of an unconstrained problem can be applied. The Lagrange function ($L^*$) can be defined as $\ln(L)$ plus the Lagrangian multipliers $\lambda(t_i)$ times the constraint function

		{\footnotesize
			\begin{equation} \label{EQ: MLE_Lagrangian}
				\begin{split}
					L^* \equiv \ln p\left( \theta \right) + \sum_{i=1}^{n} \left\{ \ln p(y(t_i)|\theta) \right. \\
					\left. + \lambda^T \cdot {\color{red} \left[ Y(t_i) - y\left( \theta,t_i \right) - \epsilon(t_i) \right] } \right\} \\ - \ln \left[ p\left( y \right) \right]
				\end{split}
		\end{equation}}

		By assumption of the relation of Equation \ref{EQ: Measurment_noise}

		{\footnotesize
			\begin{equation} \label{EQ: Noise_prob}
				p\left( y(t_i) | \theta,y_0 \right) = p\left( \epsilon(t_i) \right)
		\end{equation} }

		and specially

		{\footnotesize
			\begin{equation} \label{EQ: Noise_prob_norm}
				p\left( \epsilon(t_i) \right) = \cfrac{1}{ (2\pi)^{v/2} |\Gamma(t_i)|^{1/2} } \exp \left[ -\cfrac{1}{2}\epsilon^T(t_i) \left( \Gamma(t_i) \right)^{-1} \epsilon(t_i) \right]
		\end{equation} }
		where $|\Gamma|$ is the det$~\Gamma$, and $\Gamma$ is the covariance matrix of $\epsilon$, i.e., of the response.

		After Equation \ref{EQ: Noise_prob} is substituted into Equation \ref{EQ: MLE_Lagrangian}, and $L^*$ can be differentiated with respect to each of the estimates and $\epsilon(t_i)$. {\color{red}Based on the first order optimality conditions}, the resulting expression is equated to zero:

		{\footnotesize
			\begin{subequations} \label{EQ: MLE_Lag_diff}
				\begin{alignat}{2}
					%\cfrac{\partial L^*}{\partial y_0} &= \cfrac{\partial}{\partial y_0} \ln p(\theta,y_0) - \sum_{i=1}^{n} \lambda^T(t_i) \cfrac{\partial}{\partial y_0} y(\theta ,t_i) = 0^T \\
					\cfrac{\partial L^*}{\partial \theta} &= \cfrac{\partial}{\partial \theta} \ln p(\theta) - \sum_{i=1}^{n} \lambda^T(t_i) \cfrac{\partial}{\partial \theta} y(\theta,t_i) =0 \\
					\cfrac{\partial L^*}{\partial \epsilon(t_i)} &= \cfrac{\partial}{\partial \epsilon(t_i)} \ln p(\epsilon(t_i)) - \lambda^T(t_i) = 0^T
				\end{alignat}
		\end{subequations} }

		Substitution of Equation \ref{EQ: Noise_prob_norm} into the last Equation of \ref{EQ: MLE_Lag_diff} makes it possible to solve for $\lambda(t_i)$:

		{\footnotesize
			\begin{align}
				\lambda(t_i) &= -\left( \Gamma(t_i) \right)^{-1} \epsilon(t_i) \nonumber \\
				&= -\left( \Gamma(t_i) \right)^{-1} \left[ Y(t_i) - y(\theta,t_i) \right]
		\end{align} }

		and to eliminate $\lambda(t_i)$ from the first equation of \ref{EQ: MLE_Lag_diff}.

		For convenience we shall define a new column vector $\theta^*$ in which all the elements of $\theta$ are arranged as follows:

		{\footnotesize
			\begin{equation}
				\theta^* = \begin{bmatrix}
					\theta_{1,1} & \theta_{1,2} & \cdots & \theta_{1,v} & \theta_{2,1} & \cdots & \theta_{2,v} & \cdots & \theta_{v,v}
				\end{bmatrix}^T	
		\end{equation} }

		Finally, we assume that and $\theta^*$ are distributed by a joint normal distribution and that the prior distribution are, respectively

		{\footnotesize
			\begin{subequations} \label{EQ: MLE_Priors}
				\begin{align}
					&p\left( \theta^* \right) = \cfrac{1}{ (2\pi)^{v/2} |\Omega_{\theta^*}|^{1/2} } \exp \left[ -\cfrac{1}{2} \left( \theta^* - \theta^{*(0)} \right)^T \Omega_{\theta^*}^{-1} \left( \theta^* - \theta^{*(0)} \right) \right] 
					%&p\left( y_0 \right) = \cfrac{1}{ (2\pi)^{v/2} |\Omega_{y_0}|^{1/2} } \exp \left[ -\cfrac{1}{2} \left( y_0 - y_0^{(0)} \right)^T \Omega_{y_0}^{-1} \left( y_0 - y_0^{(0)} \right) \right] 
				\end{align}
		\end{subequations} }
		where $\Omega$'s are the respective covariance matrices for $\theta^*$ and $y_0$, and the subscript $(0)$ designates the prior estimated of $\theta^*$ and $y_0$. If we assume that $\theta^*$ and $y_0$ are independent

		%{\footnotesize
			%	\begin{equation}
				%		\ln p(\theta^*,y_0) = \ln p(\theta^*) + \ln p(y_0)
				%\end{equation} }

				Introduction of the prior distributions, Equations \ref{EQ: MLE_Priors}, plus the expression for $\lambda(t_i)$ into the first two equations of \ref{EQ: MLE_Lag_diff} gives the final equations from which the estimators of $\theta^*$ and $y_0$ can be obtained:

				{\footnotesize
					\begin{subequations} 
						\begin{flalign}
							%			&-\left( \hat{y}_0 - y_0^{(0)} \right)^T \Omega_{y_0}^{-1} + \sum_{i=1}^{n}\left[ Y(t_i) - y\left( \hat{\theta}, \hat{y}, t_i \right) \right]^T \Gamma^{-1}(t_i) \cdot \cfrac{\partial}{\partial y_0} y\left( \hat{\theta}, \hat{y}, t_i \right) = 0^T  &&\\
							&-\left( \hat{\theta}^* - \theta^{*(0)} \right)^T \Omega_{\theta^*}^{-1} + \sum_{i=1}^{n}\left[ Y(t_i) - y\left( \hat{\theta}, t_i \right) \right]^T \Gamma^{-1}(t_i) \cdot \cfrac{\partial}{\partial \theta^*} y\left( \hat{\theta}, t_i \right) = 0^T &&
						\end{flalign}
				\end{subequations} }
				where the overlay caret denotes estimated parameters.

				Note that under the assumption that the elements of $\Omega$ are essentially infinite (prior knowledge is diffuse), the elements of $\Omega^{-1}$ are zero, the equations for the maximum likelihood estimates coincide with those for the least squares estimates, and the same calculations for precision in the estimates apply.

				%For some models, these equations can be explicitly solved for $\hat{\theta}^*$ and $\hat{y}$ but in general no closed-form solution to the maximization problem is known or available, and an MLE can only be found via numerical optimization. Under the assumption of normality, the lack of prior knowledge about the parameters $\theta^0$ and known initial conditions $y_0$, the optimization problem is defined as follow:
				%
				%{\footnotesize
					%\begin{equation}
					%	\hat{\theta}_{MLE} = \arg \max_{\theta \in \Theta} L = \arg \max_{\theta \in \Theta} p(\theta|y)
					%\end{equation} }
					%
					%Intuitively, this selects the parameter values that make the observed data most probable. The specific value $\hat{\theta}$ that maximizes the likelihood function $L$ is called the maximum likelihood estimate. Further, if the function $\hat{\theta}:\mathbb{R}\rightarrow\Theta$ so defined is measurable, then it is called the maximum likelihood estimator.
					%
					%If the random samples $y(t_1)$, $y(t_2)$,...,$y(t_n)$ come from a normal distribution with unknown mean $\hat{\theta}$ and variance $\sigma^2$, then the joint probability function can be written as 
					%
					%{\footnotesize
						%\begin{equation}
						%	p(\theta|y(t_i)) = \cfrac{1}{\sqrt{2\pi\sigma}}\exp\left[-\cfrac{ \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2}{2\sigma}\right]
						%\end{equation} }
						%
						%Then the likelihood function is defined as follow:
						%
						%{\footnotesize
							%\begin{equation} 
							%	L =\prod\limits_{i=1}^n p(\theta|y(t_i)) = (2\pi\sigma)^{-n/2}\exp\left[-\cfrac{1}{2\sigma}\sum\limits_{i=1}^n \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2\right]
							%\end{equation} }
							%
							%By taking the ln of the likelihood function, the final form of the cost function is obtained:
							%
							%{\footnotesize
								%\begin{equation} 
								%	\ln L=-\cfrac{n}{2}\ln\sigma-\cfrac{n}{2}\ln(2\pi)-\cfrac{\sum\limits_{i=1}^n \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2}{2\sigma}
								%\end{equation} }
								%
								%\subsection{Maximum a posteriori estimation}
								%
								%As already discussed, the prior information about the estimated parameters can be incorporated in to the maximum likelihood estimator to form a maximum a posteriori probability (MAP) estimate. MAP can be used to obtain a point estimate of an unobserved quantity on the basis of empirical data. MAP estimation can therefore be seen as a regularization of maximum likelihood estimation. MLE and MAP can be connected by Bayes' rule:
								%
								%{\footnotesize
									%\begin{equation} 
									%	\begin{split}
										%		{\bf \hat{\theta}_{\text{MAP}}} &= \arg\max_{\sigma, \theta \in \Theta} p(\theta | y) \\
										%		&= \arg\max_{\sigma, \theta \in \Theta} \frac{p(y | \theta) p(\theta)}{p(y)} \\
										%		&= \arg\max_{\sigma, \theta \in \Theta} p(y | \theta) p(\theta) \\
										%		&= \arg\max_{\sigma, \theta \in \Theta} \ln(p(y | \theta) p(\theta)) \\
										%		&= \arg\max_{\sigma, \theta \in \Theta} \ln p(y | \theta) + \ln p(\theta)
										%	\end{split}
									%\end{equation} }
									%
									%Notice that we can get rid of the evidence term $P(y)$ because it's constant with respect to the maximization and also take the ln of the inner function because it's monotonically increasing.
									%
									%We assume the normal distribution of random samples, with identical variance and the prior distribution of $\theta$ given by $\mathcal{N}(\theta^0,\,\sigma_\theta^{2})$. The likelihood function to be maximized can be written as
									%
									%{\footnotesize
										%\begin{equation} \label{EQ: MAP_L2}
										%	\begin{split}
											%	&\arg\max_{\sigma, \theta \in \Theta} \left[ \ln \prod_{i=1}^{n} \frac{1}{\sigma\sqrt{2\pi}}e^{-\frac{ \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2 }{2\sigma^2}}
											%	+ \ln \prod_{j=0}^{p} \frac{1}{\sigma_\theta\sqrt{2\pi}}e^{-\frac{ \left( \hat{\theta} - \theta^0 \right)^2}{2\sigma_\theta^2}} \right]  \\
											%	&= \arg\max_{\sigma, \theta \in \Theta} \left[- \sum_{i=1}^{n} {\frac{\left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2}{2\sigma^2}}
											%	- \sum_{j=0}^{p} {\frac{\left( \hat{\theta} - \theta^0 \right)^2}{2\sigma_\theta^2}} \right]\\
											%	&= \arg\min_{\bf \theta} \frac{1}{2\sigma^2} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2
											%	+ \frac{\sigma^2}{\sigma_\theta^2} \sum_{j=0}^{p} \left( \hat{\theta} - \theta^0 \right)^2 \right] \\
											%	&= \arg\min_{\bf \theta} \left[ \sum_{i=1}^{n} \left( Y(t_i) - y\left( \hat{\theta}, t_i \right) \right)^2 + \lambda_{L2} \sum_{j=0}^{p} \left( \hat{\theta} - \theta^0 \right)^2 \right]
											%	\end{split}
										%\end{equation} }
										%
										%Notice that we dropped some of the constants (with respect to $\theta$) and factored to simplify the expression. You can see this is the same expression as Equation \ref{EQ: MAP_L2} (L2 Regularization) with $\lambda_{L2}$. We can adjust the amount of regularization we want by changing $\lambda_{L2}$. Equivalently, we can adjust how much we want to weight the priors carry on the coefficients ($\theta$). If we have a very small variance (large $\lambda_{L2}$) then the coefficients will be very close to 0; if we have a large variance (small $\lambda_{L2}$) then the coefficients will not be affected much (similar to as if we didn't have any regularization).

	\fi

\end{document}