### 2023-04-22 SFE-article/Sections/Experiments.tex

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\label{CH: Experiments}

	The close the optimization problem given by a set of equations \ref{EQ: Optimization_formulation_MLE}, it is required to know the dataset $Y(t)$. In this case, the dataset has been obtained by performing the extraction of caraway oil from caraway seeds. The caraway seeds have been obtained from the company Caraway Finland\todo{Should be mentioned?} during the 2022 growing season. The caraway seeds have been pre-treated with Retsch SM 300 cutting mill to decrease the particle size to 1mm and break the outer shell of the seeds. Later, the moisture content was analysed to ensure that it does not have a negative effect on the extraction process. An Infrared Moisture Analyser was used to investigate the sample. The moisture content after grinding was 4.83\%. The grind material density was measured by pycnometer and its result can be found in appendix \ref{CH: Solid_Density_Measurment}. Moreover, the porosity was calculated to be 0.5, as explained in appendix \ref{CH: Porosity}.

	The unit used for the experimental work is a 10-liter extractor \todo{Should Biorukki be mentioned?} with an inner diameter of around 15 cm and 60 cm height \todo{Double check dimension}. Four experiments have been performed in that unit at the following operating conditions: $40^\circ C / 200$ bar, $50^\circ C / 200$ bar, $40^\circ C / 300$ bar and $50^\circ C / 300$ bar. 1 kg (or 1.6 liters) of grind material was weighed and placed in the extraction chamber. Due to technical limitations, the amount of solid material for extraction was selected to be 1 kg per batch, which is not sufficient to fill the whole chamber.

	After the material was loaded into the chamber, the extractor is pre-heated to desired temperature. The next step is to close the outlet line coming out of the extractor and fill the extraction chamber with $CO_2$. The $CO_2$ is pumped and compressed until the operating pressure is achieved. When the operating temperature and pressure are obtained, the outlet in is open to allow the solvent to flow through the system. The solvent extracts essential oils from the fixed bed of solid material and the mixture $CO_2$-oil flow from the extractor to the separator. The separator was operating at $50^\circ C$ and $50$ bar. At this condition, the gasified $CO_2$ leaves from the top, while oil stays in the liquid form and is collected at the bottom of the separator. Every 5 minutes, the oil is drained from the separator, and its weight is measured. The amount of oil collected during a batch is an output of the experiment and it is used as data for parameter estimation. The extraction time for each batch was 150 minutes, counted from the opening of the extractor outlet port.

	The results of experiments are presented in the table \ref{tab: Yield_data}.

	\begin{table}[tbh!]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{c|cccc}
			& $40 ^\circ C / 200$ bar & $50 ^\circ C / 200$ bar & $40 ^\circ C / 300$ bar & $50 ^\circ C / 300$ bar \\ 
			Time min & Mass g                  & Mass g                  & Mass g                  & Mass g                  \\ \hline
			0        & 0.0                     & 0.0                     & 0.0                     & 0.0                     \\
			5        & 1.7                     & 1.8                     & 1.1                     & 1.2                     \\
			10       & 5.7                     & 4.7                     & 4.7                     & 4.5                     \\
			15       & 10.9                    & 9.5                     & 10.8                    & 11.3                    \\
			20       & 17.8                    & 15.1                    & 20.6                    & 20.6                    \\
			25       & 24.9                    & 21.3                    & 32.1                    & 30.6                    \\
			30       & 30.3                    & 26.4                    & 40.5                    & 37.8                    \\
			35       & 34.0                    & 29.8                    & 46.7                    & 43.3                    \\
			40       & 37.5                    & 33.1                    & 50.5                    & 47.5                    \\
			45       & 40.1                    & 35.6                    & 53.1                    & 50.8                    \\
			50       & 43.3                    & 38.7                    & 55.5                    & 53.7                    \\
			55       & 45.5                    & 41.5                    & 57.6                    & 56.5                    \\
			60       & 47.6                    & 43.5                    & 59.3                    & 58.1                    \\
			65       & 49.2                    & 45.9                    & 60.9                    & 59.6                    \\
			70       & 50.8                    & 48.3                    & 62.2                    & 61.1                    \\
			75       & 52.6                    & 50.8                    & 63.7                    & 62.7                    \\
			80       & 54.2                    & 52.9                    & 64.9                    & 64.4                    \\
			85       & 55.1                    & 55.3                    & 66.6                    & 65.8                    \\
			90       & 56.2                    & 57.2                    & 67.6                    & 66.7                    \\
			95       & 57.3                    & 58.3                    & 68.9                    & 67.4                    \\
			100      & 58.3                    & 59.4                    & 69.7                    & 68.2                    \\
			105      & 59.4                    & 60.4                    & 70.6                    & 69.1                    \\
			110      & 60.2                    & 61.6                    & 71.2                    & 70.1                    \\
			115      & 61.4                    & 62.7                    & 71.9                    & 70.9                    \\
			120      & 62.5                    & 63.9                    & 72.3                    & 71.6                    \\
			125      & 63.3                    & 64.7                    & 72.8                    & 72.3                    \\
			130      & 64.0                    & 65.4                    & 73.2                    & 72.9                    \\
			135      & 64.8                    & 65.9                    & 73.6                    & 73.5                    \\
			140      & 65.5                    & 66.4                    & 74                      & 74                      \\
			145      & 66.2                    & 67.1                    & 74.3                    & 74.5                    \\
			150      & 66.8                    & 67.7                    & 74.6                    & 74.9                   
		\end{tabular} }
	\caption{Yield data}
	\label{tab: Yield_data}
	\end{table}

\end{document}