### 2023-05-07 SFE-article/Sections/Profiles.tex

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	The time evolution of concentration profiles along the extractor for each dataset is presented on Figure \ref{fig: estimation_results_profiles}. The concentration profiles use the parameters obtained from the parameter estimation. The plots on the left hand-side show the fluid phase concentration, while the plots on the right-hand side present the concentration in the solid phase.

	\begin{figure}[!h]
		\centering
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 1.4cm 1.9cm 11cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P200.pdf}
			\caption{Experiment at $40^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 1.4cm 1.9cm 11cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P200.pdf}
			\caption{Experiment at $50^\circ C$ and $200$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 1.4cm 1.9cm 11cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T40_P300_org.pdf}
			\caption{Experiment at $40^\circ C$ and $300$ bar}
		\end{subfigure}
		\begin{subfigure}[b]{\columnwidth}
			\centering
			\includegraphics[trim = 2cm 1.4cm 1.9cm 11cm,clip,width=\textwidth]{/Results_estimation/Fitting_LUKE_T50_P300_org.pdf}
			\caption{Experiment at $50^\circ C$ and $300$ bar}
		\end{subfigure}
		\caption{Concentration profiles after solving the parameter estimation problem}
		\label{fig: estimation_results_profiles}
	\end{figure}

\end{document}