### 2023-05-07 SFE-article/Sections/Saturation.tex

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\subsubsection{Saturation Concentration}

	The above model does not take into account the saturation of fluid, which can be introduced by multiplying the gradient by a function $\gamma(c_f)$ (equation \ref{EQ: C_sat_function}). $\gamma(c_f)$ describe the reverse logistic function, which is equal to unity below the $c_{sat}$, the saturation concentration, and equal to zero, above the $c_{sat}$. The $\gamma(c_f)$ for $c_{sat}=5$ is shown on figure \ref{fig: Gamma_function}.

	{\footnotesize
		\begin{equation}
			\gamma(c_f) = \cfrac{1}{1+\exp ( -k_{sat} \left( c_f - c_{sat} \right) ) }
			\label{EQ: C_sat_function}
			\end{equation}
	}

	where $k_{sat}$ is the growth rate, and it is defined as $k_{sat} = 4$. The growth rate corresponds for the shape of the $\gamma(c_f)$ function. The bigger value of the $k_{sat}$ the steeper the function is. In the limit when $ k_{sat}\rightarrow+\infty$ become a step function. If the growth rate is to high, the gradient of the $\gamma$ function become high, or goes to infinity, which might cause difficulties for a gradient-based optimizer.

	\begin{figure}[h!]
		\centering
			\begin{tikzpicture}

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 10,
					ymin = -0.5, ymax = 1.5,
					xlabel = {$c_f[kg/m^3]$},
					ylabel = {$\gamma(c_f)[-]$}]
					\addplot[
					domain = 0:10,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 4 .* ( x - 5 ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

					\addlegendentry{$\gamma(c_f)$}
					\addlegendentry{$c_{sat}$}
				\end{axis}

			\end{tikzpicture}
			\caption{$\gamma(c_f)$ function under assumption of $c_{sat}=5 [kg/m^3]$}
			\label{fig: Gamma_function}
	\end{figure}

	The final form of the extraction kinetic equation is given by equation \ref{Model_kinetic}.

	{\footnotesize
		\begin{equation}
			\label{Model_kinetic}
				{\color{blue}r_e}(t,z) = -\cfrac{{\color{orange}D_i}({\color{blue}T}(t,z), {\color{red}P}(t))}{{\color{magenta} \mu l^2} } \gamma(c_f) \left({\color{blue}c_s}(t,z) - \cfrac{{\color{magenta}\rho_s}}{{\color{orange}k_m}({\color{blue}T}(t,z)){\color{orange}\rho}({\color{blue}T}(t,z),{\color{red}P}(t))}  {\color{blue}c_f}(t,z) \right)
		\end{equation} }

\end{document}