### 2023-06-21 SFE-article/Sections/Saturation.tex

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\subsubsection{Concentration limit}

	To mimic that not all solute might be obtained at specific operating conditions, the coefficient ${\color{orange} \gamma}$ is introduced to limit the available oil. The ${\color{orange} \gamma}({\color{blue} c_s} (t,z))$ describe the reverse logistic function, which is equal to unity above the ${\color{magenta} c_{lim}}$, the saturation concentration, and equal to zero, below the ${\color{magenta} c_{lim}}$. The ${\color{orange} \gamma}$ parameter multiply the solute concentration term in solid phase in the concentration gradient. As the concentration of solute decrease, it can reach the limiting value ${\color{magenta} c_{lim}}$ when the ${\color{orange} \gamma}$ goes to zero and stop the extraction from the solid.

	{\footnotesize
		\begin{equation}
			{\color{orange} \gamma}({\color{blue} c_s} (t,z)) = \cfrac{1}{1+\exp \left( -{\color{magenta} k_{lim}} \left( {\color{blue} c_s} (t,z) - {\color{magenta} c_{lim}} \right) \right) }
			\label{EQ: C_sat_function}
			\end{equation}
	}

	where ${\color{magenta} k_{lim}}$ describe the curvature of the $\gamma$ function, and it is defined as ${\color{magenta} k_{lim}} = 1$. %In the limit when $ {\color{magenta} k_{lim}}\rightarrow+\infty$ become a step function. If the ${\color{magenta} k_{lim}}$ is to high, the gradient of the ${\color{orange} \gamma}$ function become high, or goes to infinity, which might cause difficulties for a gradient-based optimizer. 
	The ${\color{orange} \gamma}({\color{blue} c_s} (t,z))$ gamma function is visualised on \ref{fig: Gamma_function}.

	\begin{figure}[h!]
		\centering
			\begin{tikzpicture}

				\begin{axis}[
					legend style={draw=none},
					xmin = 0, xmax = 100,
					ymin = -0.5, ymax = 1.5,
					xlabel = {${\color{blue} c_s}(t,z)[kg/m^3]$},
					ylabel = {${\color{orange} \gamma}({\color{blue} c_s}(t,z))[-]$}]
					\addplot[
					domain = 0:100,
					samples = 100,
					smooth,
					thick,
					blue,
					] { 1 / (1 + exp( 1 * ( 50 - x ) ) ) };
					\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(50,-0.5)(50,1.5)};

					\addlegendentry{${\color{orange} \gamma}({\color{blue} c_s} (t,z))$}
					\addlegendentry{${\color{magenta} c_{lim}}$}
				\end{axis}

			\end{tikzpicture}
			\caption{${\color{orange} \gamma}({\color{blue} c_s} (t,z))$ function under assumption of ${\color{magenta} c_{lim}}=50~[kg/m^3]$}
			\label{fig: Gamma_function}
	\end{figure}

	The final form of the extraction kinetic equation is given by equation \ref{Model_kinetic}.

	{\scriptsize
		\begin{equation}
			\label{Model_kinetic}
				{\color{blue}r_e}(t,z) = -\cfrac{{\color{orange}D_i}({\color{blue}T}(t,z), {\color{red}P}(t))}{{\color{magenta} \mu l^2} } \left({\color{blue}{\color{blue} c_s} }(t,z) {\color{orange} \gamma}({\color{blue} c_s} (t,z)) - \cfrac{{\color{magenta}\rho_s}}{{\color{orange}k_m}({\color{blue}T}(t,z)){\color{orange}\rho}({\color{blue}T}(t,z),{\color{red}P}(t))}  {\color{blue}c_f}(t,z) \right)
		\end{equation} }

\end{document}