### 2023-10-05 SFE_article_Parameter_Estimation/Sections/Qubic_EOS.tex
05.10

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\label{CH: Thermodynamic_details}

	subsubsection{Equation of state and properties of the fluid phase} \label{subsubsec: Equation of state}

	A cubic equation of state is a mathematical model used in the field of thermodynamics and fluid dynamics to describe the behavior of real gases and liquids. It is named "cubic" because it is characterized by a third-degree polynomial equation that relates the pressure, volume, and temperature of a substance. These equations are essential tools for understanding the phase behavior, properties, and thermodynamic processes of real substances, making them invaluable in various engineering and scientific applications. The cubic equation of state accounts for the deviations from ideal gas behavior, which is particularly important at high pressures and low temperatures, where real gases show significant departures from ideal behavior.

	The general cubic equation of state is typically written in the following form:

	{\footnotesize
	\begin{equation}
		P = \frac{RT}{v_m-b} - \frac{a\alpha}{v_m^2 - ubv_m + wb^2}
	\end{equation}
	}

	Where P is the pressure of the substance, $v_m$ is the molar volume (volume per mole) of the substance, T is the absolute temperature of the substance., u and w are integers that differ rom an equation to another, R is the universal gas constant, a and b are substance-specific parameters known as the Van der Waals constants and $\omega$ is an acentric factor.

	The Van der Waals constants, a and b, are empirical values that depend on the specific substance being modeled. These constants account for the molecular interactions (attractive forces represented by a) and the finite size of gas molecules (excluded volume represented by b). For different substances, a and b will have different values based on experimental data.

	Several variations of the cubic equation of state exist, including the Van der Waals equation and the Redlich-Kwong equation, each with its own set of parameters and assumptions. Tables \ref{tab:Popular_Cubic_EoS} show parameters for popular cubic EoS.

	\begin{table}[h!]
		\centering
		\adjustbox{width=\columnwidth}{%
			\begin{tabular}{|c| c c c c|} 
				\hline
				EoS & u & w & a & b\\
				\hline
				van der Waals  & 0 & 0 & $\frac{27}{64} \frac{R^2T_c^2}{P_c}$ & $\frac{RT_c}{8P_c}$ \\
				Redlich and Kwong & 1 & 0 & $0.42748 \frac{R^2T_c^{2.5}}{P_c}$ & $\frac{0.08664RT_c}{P_c}$ \\
				Soave & 1 & 0 & $0.42748 \frac{R^2T_c^2}{P_c}$ & $\frac{0.08664RT_c}{P_c}$ \\
				Peng and Robinson \cite{Peng1976} & 2 & -1 & $0.45724 \frac{R^2T_c^2}{P_c}$ & $\frac{0.07780T_c}{P_c}$\\
				\hline
		\end{tabular} }
		\caption{Parameters for Popular Cubic EoS}
		\label{tab:Popular_Cubic_EoS}
	\end{table}

	\begin{table}[h!]
		\centering
		\adjustbox{width=\columnwidth}{%
			\begin{tabular}{|c| c c|} 
				\hline
				EoS & $\alpha$ & f($\omega$)\\
				\hline
				van der Waals  & - & - \\
				Redlich and Kwong & $\frac{1}{\sqrt{T_r}}$ & - \\
				Soave & $\left[ 1 + f(\omega) \left( 1-\sqrt{T_r} \right) \right]^2$ & 0.48+1.574$\omega$-0.176$\omega^2$\\
				Peng and Robinson \cite{Peng1976} & $\left[ 1 + f(\omega) \left( 1-\sqrt{T_r} \right) \right]^2$ & 0.37464+1.54226$\omega$-0.26992$\omega^2$ \\
				\hline
		\end{tabular} }
		\caption{Parameters for Popular Cubic EoS}
		\label{tab:Popular_Cubic_EoS_alpha}
	\end{table}

	The general cubic equation of state can be written in polynomial as presented by Equation \ref{EQ:Compressibility_Polynomial}. In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. The gas-liquid mixture is present in the two-phase region, and two roots are found. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

	{\footnotesize
		\begin{equation}
			\label{EQ:Compressibility_Polynomial}
			Z^3 - (1+B-uB)Z^2+(A+wB^2-uB-uB^2)Z - AB - wB^2 - wB^3 = 0
	\end{equation} }

	where $A=\frac{a\alpha P}{R^2T^2}$ and $B=\frac{bP}{RT}$.

	If the Peng-Robinson equation of state \citet{Peng1976} is used, the polynomial equation becomes

	{\footnotesize
	\begin{equation}
		\label{EQ:Peng_Robinson_Polynomial}
		Z^3 - (1-B)Z^2+(A-2B-3B^2)Z -(AB-B^2-B^3) = 0
	\end{equation} }

	For an ideal gas the compressibility factor is $Z = 1$ per definition. In many real world applications requirements for accuracy demand that deviations from ideal gas behaviour, i.e., real gas behaviour, be taken into account. The value of $Z$ generally increases with pressure and decreases with temperature. At high pressures molecules are colliding more often. This allows repulsive forces between molecules to have a noticeable effect, making the molar volume of the real gas ($v_m$) greater than the molar volume of the corresponding ideal gas $\left( \left(v_m\right)_{ideal~gas} = \frac{RT}{P} \right)$, which causes Z to exceed one. When pressures are lower, the molecules are free to move. In this case attractive forces dominate, making Z < 1. The closer the gas is to its critical point or its boiling point, the more Z deviates from the ideal case.

	Numerical methods such as Newton-Raphson can be used to solve the polynomial equation to obtain the compressibility ${\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ at given temperature and pressure. Alternatively, the Cardano formula can be used to get the closed form solution of the Equation \ref{EQ:Compressibility_Polynomial}. Further details regarding the Cardano formula can be found in Appendix \ref{CH: Cardano}.

	\subsubsection{Density of the fluid phase} \label{subsubsec: Fluid density}

	The density of the fluid can be calculated from the real gas equation $\rho = \frac{P}{RTZ} \frac{1}{m_{CO2}}$. The local temperature can be obtain from the time evolution of governing equations, the pressure is consider to be constant along the system to be a know at any time. The local compressibility factor can be also computed if the local temperature and pressure are know. To obtain the density in terms of mass, molar mass of the solvent is taken into account.

	\subsubsection{Heat capacity of the fluid phase} \label{subsubsec: Fluid heat capacity}
	%	\noindent\textbf{Heat capacity of the fluid phase} \\

	%The specific heat $C_p^{\text{F}}$ can be calculated from the equation of state, again under the assumption that the fluid phase consists of pure carbon dioxide and that the specific heat of real fluids can be calculated from an ideal contribution plus a residual term \citet{Pratt2001}. The main steps in the derivation of ${\color{black}C_p}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ are presented below: 
	Following the work of \citet{Pratt2001}, it is assumed that the specific heat can be obtained as a sum of ideal gas contribution and a residual correction for non-ideal behaviour, as given by Equations \ref{eq: Cv_CO2} and \ref{eq: Cp_CO2}.

	{\footnotesize
		\begin{subequations}\begin{align}
				{\color{black}C_v}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) & = {\color{black}C_v^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right) + {\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) \label{eq: Cv_CO2}; \\ 
				{\color{black}C_p}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) & = \underbrace{{\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right)}_{\text{Eq. } \eqref{eq: Cp_I}} + \underbrace{{\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}_{\text{Eq. } \eqref{eq: Cp_R}}  \label{eq: Cp_CO2}.
	\end{align}\end{subequations} }

	${\color{black}C_v}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ and ${\color{black}C_p}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ are the specific heat of $CO_2$ at constant volume and pressure, respectively.

	The ideal-gas contribution ${\color{black}C_v^{\text{I}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$ and ${\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)$, can be obtained based on heat-capacity data applicable to gases at low pressure. For $CO_2$, the ideal gas contribution to the specific heat at constant ${\color{black}P}(t)$, as function of ${\color{black}T}(t,z)$ is given by,

	{\footnotesize
		\begin{equation}
			{\color{black}C_p^{\text{I}}}\left({\color{black}T}(t,z),P(t)\right) = {\color{black}C_{P0}} + {\color{black}C_{P1}}T(t,z) + {\color{black}C_{P2}}{\color{black}T}^2(t,z) + {\color{black}C_{P3}}{\color{black}T}^3(t,z) \label{eq: Cp_I}
	\end{equation} }

	where the coefficients of the expansion are ${\color{black}C_{P0}} = 22.26$, ${\color{black}C_{P1}} = 5.981 \times 10^{-2}$, ${\color{black}C_{P2}} = -3.501 \times 10^{-5}$, and ${\color{black}C_{P3}} = 7.469 \times 10^{-9}$, as given by \citet{Kyle1999}.

	The correction term ${\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z){\color{black}P}(t)\right)$ at constant pressure ${\color{black}P}(t)$, can be defined by introducing  ${\color{black}C_p^{\text{I}}}({\color{black}T}(t,z)) - {\color{black}C_v^{\text{I}}}({\color{black}T}(t,z)) = {\color{black}R}$ and using the definition of ${\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z){\color{black}P}(t)\right)$. The general form of the correction term is given by Equation \ref{eq: Cp_R}

	{\footnotesize	
		\begin{align}
			&{\color{black}C_p^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) = \underbrace{{\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}_{\text{Eq. }\eqref{eq: CvR}} \nonumber \\
			&+ {\color{black}T}(t,z) \underbrace{\left(\cfrac{\partial {\color{black}P}(t)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)}}_{\text{Eq. } \eqref{eq: dPdT}} \underbrace{\left(\cfrac{\partial {\color{black}v_m}\left({\color{black}T}(t,z),{\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)}}_{\text{Eq. } \eqref{eq: dVdT}} - {\color{black}R}. \label{eq: Cp_R} 
	\end{align} }

	%	The braced terms are obtained from the chosen equation of state and the compressibility definition ${\color{black}P}(t) {\color{black}v_m}({\color{black}T}(t,z),{\color{black}P}(t)) = {\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right){\color{black}R}{\color{black}T}(t,z)$. \\

	For the partial derivative of the volume with respect to temperature ${\color{black}T}(t,z)$ at constant pressure ${\color{black}P}(t)$, we have

	{\footnotesize
		\begin{align}
			\left( \cfrac{\partial {\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}} \right)_{{\color{black}P}(t)} &= \cfrac{{\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right) {\color{black}R}}{{\color{black}P}(t)} \nonumber \\
			&+ \cfrac{{\color{black}R}{\color{black}T}(t,z)}{{\color{black}P}(t)} \underbrace{\left( \cfrac{\partial {\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}} \right)_{{\color{black}P}(t)}}_{\text{Eq. }\eqref{eq: dZdT}} \label{eq: dVdT}
	\end{align} }

	with partial derivative of the compressibility factor with respect to temperature ${\color{black}T}(t,z)$ at constant pressure ${\color{black}P}(t)$

	{\footnotesize
		\begin{equation}
			\left(\cfrac{\partial{\color{black}Z}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)} = \left(\cfrac{\partial \dfrac{{\color{black}P}(t) {\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}{{\color{black}R}{\color{black}T}(t,z)}}{\partial {\color{black}T}}\right)_{{\color{black}P}(t)} \label{eq: dZdT}
	\end{equation} }

	Similarly, for the partial derivative of the pressure with respect to temperature at constant volume is :

	{\footnotesize
		\begin{equation}
			\left(\cfrac{\partial {\color{black}P}(t)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)} = \left( \cfrac{\partial{\dfrac{{\color{black}Z}{\color{black}R}{\color{black}T}(t,z)}{{\color{black}v_m}\left({\color{black}T}(t,z), {\color{black}P}(t)\right)}}}{\partial{\color{black}T}}\right)_{{\color{black}v_m}(t,z)} \label{eq: dPdT}
	\end{equation} }

	The residual specific heat at constant volume is obtained, by definition, by using the residual internal energy

	{\footnotesize
		\begin{equation}
			{\color{black}C_v^{\text{R}}}\left({\color{black}T}(t,z),{\color{black}P}(t)\right) = \left(\cfrac{\partial{\color{black}U^{\text{R}}}\left({\color{black}T}(t,z),{\color{black}P}(t)\right)}{\partial {\color{black}T}}\right)_{{\color{black}v_m}(t,z)} \label{eq: CvR}
	\end{equation} }

	\subsubsection{Departure functions for enthalpy calculations} \label{CH:Enthalpy}

	In thermodynamics, a departure function is a concept used to calculate the difference between a real fluid's thermodynamic properties and those of an ideal gas, given a specific temperature and pressure. Common departure functions include those for enthalpy, entropy, and internal energy. These functions are used to calculate extensive properties, which are properties computed as a difference between two states.

	To evaluate the enthalpy change between two points, $h(V_1,T_1)$ and $h(V_2,T_2)$, the enthalpy departure function between the initial volume $V_1$ and infinite volume at temperature $T_1$ is calculated. Then it is added to that the ideal gas enthalpy change due to the temperature change from $T_1$ to $T_2$, and finally subtract the departure function value between the final volume $V_2$ and infinite volume.

	Departure functions are computed by integrating a function that depends on an equation of state and its derivative. The general form of the enthalpy equation is given by:

	{\footnotesize
		\begin{equation}
			\frac{h^{id}-h}{RT} =\int_{v_m}^{\infty }\left(T\left({\frac{\partial Z}{\partial T}}\right)_{v_m}\right){\frac{dv_m}{v_m}} + 1-Z
		\end{equation}
	}

	Here, $h^{id}$ represents the enthalpy of an ideal gas, $h$ is the enthalpy of a real fluid, $R$ is the universal gas constant, $T$ is temperature, $v_m$ is the molar volume, and $Z$ is the compressibility factor.

	The integral in the equation is evaluated over the range of molar volumes from $v_m$ to infinity. The integral includes a term that depends on the derivative of the compressibility factor with respect to temperature, evaluated at the molar volume $v_m$. Finally, the term $1-Z$ is added to account for the deviation of the fluid's properties from those of an ideal gas.

	The Peng–Robinson equation of state relates the three interdependent state properties pressure $P$, temperature $T$, and molar volume $v_m$. From the state properties ($P$, $v_m$, $T$), one may compute the departure function for enthalpy per mole (denoted $h$) as presented by \citet{Gmehling2019} or \citet{Elliott2011}:

	{\footnotesize
		\begin{equation}
			h-h^{\mathrm {id} }=RT_c\left[T_{r}(Z-1)-2.078(1+\kappa ){\sqrt {\alpha }}\ln \left({\frac {Z+2.414B}{Z-0.414B}}\right)\right]
		\end{equation}
	}

\end{document}