### 2023-10-11 SFE_article_Sensitivity_Analysis/Article_Model_Parameters.tex
11.10

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. Use 1, 2 and 3. TO be removed later
%\paperwidth=\dimexpr \paperwidth + 6cm\relax
%\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
%\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{graphicx,float}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}

\usepackage{csvsimple}

% TODO package
\usepackage[bordercolor=gray!20,backgroundcolor=blue!10,linecolor=black,textsize=footnotesize,textwidth=1in]{todonotes}
\setlength{\marginparwidth}{1in}
% \usepackage[utf8]{inputenc}
% \usepackage[nomath]{lmodern}

% Margin and formatting specifications
%\usepackage[authoryear]{natbib}
\usepackage[sort]{natbib}
\setcitestyle{square,numbers}

 %\bibliographystyle{cas-model2-names}

\usepackage{setspace}
\usepackage{subfiles} % Best loaded last in the preamble

% \usepackage[authoryear,longnamesfirst]{natbib}

% Math packages
\usepackage{amsmath, amsthm, amssymb, amsfonts, bm, nccmath, mathdots, mathtools, bigints, ulem}

\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{shapes.geometric,angles,quotes,calc}

\usepackage{placeins}

\usepackage[final]{pdfpages}

% --------------------------------------------------------------------
% Packages Configurations
\usepackage{enumitem}
% --------------------------------------------------------------------
% (General) General configurations and fixes
\AtBeginDocument{\setlength{\FullWidth}{\textwidth}}	% Solves els-cas caption positioning issue
\setlength{\parindent}{20pt}
%\doublespacing
% --------------------------------------------------------------------
% Other Definitions
% --------------------------------------------------------------------
\graphicspath{{Figures/}}
% --------------------------------------------------------------------
% Environments
% --------------------------------------------------------------------
% ...

% --------------------------------------------------------------------
% Commands
% --------------------------------------------------------------------

% ==============================================================
% ========================== DOCUMENT ==========================
% ==============================================================
\begin{document} 
%  --------------------------------------------------------------------

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Sensitivity Analysis}                      
\shorttitle{Sensitivity Analysis}

\shortauthors{OS, PO}

\author[1]{Oliwer Sliczniuk}[orcid=0000-0003-2593-5956]
\ead{oliwer.sliczniuk@aalto.fi}
\cormark[1]
\credit{a}

\author[1]{Pekka Oinas}[orcid=0000-0002-0183-5558]
\credit{b}

%\author[1]{Francesco Corona}[orcid=0000-0002-3615-1359]
%\credit{c}

\address[1]{Aalto University, School of Chemical Engineering, Espoo, 02150, Finland}
%\address[2]{2}

\cortext[cor1]{Corresponding author}

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from caraway seeds. The extraction was performed in a partially filled extractor with a fixed bed operated under multiple operating conditions. A distributed-parameter model describes the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The local sensitivity analysis investigates the influence of infinitely small changes in the inlet temperature, pressure, and flow rate on the extraction yield.

\end{abstract}

\begin{keywords}
Supercritical extraction \sep Sensitivity analysis \sep Mathematical modelling
\end{keywords}

% ===================================================
% TITLE
% ===================================================
\maketitle

% ===================================================
% Section: Introduction
% ===================================================\section{Introduction}

\section{Introduction}
%\subfile{Sections/introduction}
\subfile{Sections/introduction_imp}

\section{Materials and methods} \label{CH: Materials and methods}

\subsection{Supercritical fluids} \label{CH: Thermodynamic}
%\subfile{Sections/Thermo}
\subfile{Sections/Thermo_imp}

%\clearpage
%\subfile{Sections/Materials_and_methods}

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
\subfile{Sections/Gouverning_equation}

\subsection{Extraction model} \label{CH: Extraction_model}
\subfile{Sections/Model}

%\newpage
%\section{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}
%\clearpage

\subsection{Sensitivity Analysis} \label{CH: Sensitivity_Analysis}
\subfile{Sections/Sensitivity_Analysis}

%\subsection{Experimental work}
%\subfile{Sections/Experiments}

\section{Results}
\subfile{Sections/Results_Sensitivity}

\section{Conclusions} \label{CH: Conclusion}

The sensitivity analysis is a tool to understand the parametric dependence of dynamic behavior and its suitability in inferring mechanisms of the dynamic behavior of the analyzed system. The mathematical formulation involves derivative-based local sensitivity analysis of the solution of an inverse problem with respect to perturbations of parameters. The local method is based only on considering all parameters as constants (those independent of time or location), and then the sensitivity coefficients are estimated by solving a variational system simultaneously with the original system. The sensitivity equations can be obtained in various ways, but in the case of this work, the automatic differentiation technique was implemented to obtain the sensitivity equations. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can yield valuable information for guiding the design of experiments, the process modeling, or the model reduction. The local sensitivity analysis techniques consider the sensitivities at only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear system.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity equations evaluate the influence of control variables such as flow rate, pressure, and inlet temperature on the state-space consisting of the concentration of the solute in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance. This information can be utilized to identify which control variables influence the extraction yield the most. The controls with high sensitivities on the extraction can be used to investigated to find optimal operating conditions from economic point of view.

% ===================================================
% Bibliography
% ===================================================
%% Loading bibliography style file
\clearpage
%\bibliographystyle{model1-num-names}
\bibliographystyle{unsrtnat}
\bibliography{mybibfile}

%\clearpage \appendix \label{appendix}
%\section{Appendix} 
%\subsection{Thermodynamic}
%\subfile{Sections/Qubic_EOS}

%\subsection{Governing equations}
%\subfile{Sections/Gouverning_equation_derivation}

%\subsection{Low Mach number expansion} \label{CH:Low_Mach_chapter}
%\subfile{Sections/Low_mach_number_expansion}
%\subfile{Sections/Low_mach_number_expansion_imp}

%\subsection{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}

%\subsection{Cardano's Formula} \label{CH: Cardano}
%\subfile{Sections/Cardano}

%\subsection{Initial and boundary conditions} \label{CH: IC_BC}
%\subfile{Sections/IC_BC}

%\subsection{Maximum Likelihood} \label{CH: ML}
%\subfile{Sections/Likelihood}

%\subsection{Solid density measurement} \label{CH: Solid_Density_Measurment}

%\begin{figure}[!h]
%	\centering 
%	\includegraphics[trim=2cm 6cm 4cm 0cm, clip,width=\columnwidth]{Sections/ultraReportT5.pdf}
%	\caption{The result of solid density measurement}
%\end{figure}

%{\footnotesize
%	\begin{equation*}
%		\rho_s^{ave} = \frac{1.2585+1.2582+1.2561+1.2546+1.2555}{5} = 1.25658 [g/cc]
%	\end{equation*} }

%\subsection{Porosity Calculations} \label{CH: Porosity}
%\subfile{Sections/Porosity}

%\subsection{Concentration profiles} \label{CH: Profiles}
%\subfile{Sections/Profiles}

%\subsection{Solution of the parameter estimation starting from another initial guess} \label{CH: Profiles_2}
%\subfile{Sections/Parameter_estimation_solution_from_another_IG}

\end{document}