### 2024-05-08 SFE_article_Sensitivity_Analysis_short - Chamomile/Sections/Sensitivity_Pressure.tex

\documentclass[../Article_Sensitivity_Analsysis.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	\subsection{Pressure}

	As discussed in Chapter \ref{CH:Governing_equations_chapter}, a small pressure wave propagates at the speed of sound relative to the flow. If the flow velocity is relatively low, all pressure changes are hydrodynamic (resulting from velocity motion) rather than thermodynamic. The Low Mach-number assumption leads to instant propagation of the thermodynamic pressure throughout the system. This assumption allows considering a single pressure value for the entire system, as all changes occur simultaneously within the machine. Figure \ref{fig:Sensitivty_P_P} illustrates a step function representing the pressure change.

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=0.9\columnwidth]{/Results_sensitivity/P_P.png}
		\caption{The effect of $P$ change on $P$ in the system}
		\label{fig:Sensitivty_P_P}
	\end{figure}

	According to Equation \ref{EQ:Enthalpy_equation}, the pressure change directly affects the quantity $h \times \rho$ through $\frac{\partial (P(t) A_f)}{\partial t}$, leading to the step change along the whole system, as presented in Figure \ref{fig:Sensitivty_P_H}. The uniform response across the entire extraction column length and time, represented by the homogeneously dark red colour, indicates that the entire system experiences an immediate and uniform change in enthalpy density in response to pressure changes.

	The pressure change influences the temperature of the fluid in the middle of the computational domain, but the values on boundaries can be defined in multiple ways. Applying the Dirichlet boundary conditions assumes a fixed temperature value at the inlet, which leads to a possible thermal gradient propagating along the system. In contrast, Neumann boundary conditions would dictate that the heat flux at the boundaries be held constant. In this work, the Neumann boundary conditions have been applied, which leads to a uniform response and ensures the temperatures at the inlet and in the middle of the extractor are the same.

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=0.9\columnwidth]{/Results_sensitivity/H_P.png}
		\caption{The effect of $P$ change on $(h \times \rho)$ in the system}
		\label{fig:Sensitivty_P_H}
	\end{figure}

	Figure \ref{fig:Sensitivty_P_CS} shows the sensitivity of the solute concentration in the solid phase with respect to changes in pressure throughout a fixed bed in a supercritical fluid extraction process. As discussed in Chapter \ref{CH: Continuity}, the velocity of a fluid is inversely related to its density, which suggests that with higher fluid density, the velocity decreases. This leads to an extended residence time, ergo, a longer interaction between the solute and the solvent.

	Initially, the extraction process is in the kinetically-controlled regime, where the concentration gradient is high, and the limiting factor is the solute solubility. As explained in ({\color{red}article 1}), the system is considered to be far from the saturation, which can explain low system response at the beginning of the process.

	When the concentration gradient diminishes, and the extraction moves from the kinetic-controlled to the diffusion-controlled regime, the system response becomes more profound. ({\color{red}article 1}) obtained a linear correlation for $D_i^R$, which is incorporated in this model. $D_i^R$ increases together with fluid density, which explains negative sensitivities. The negative sign can be interpreted as a faster loss of a solute from a solid phase, which corresponds to enhanced mass transfer.

	Over time, the amount of solute becomes a limiting factor and the pressure change has a lower effect on the system and the sensitivities increase. Eventually, sensitivities asymptotically approach zero. The solute concentration in the solid phase has been reduced, and further changes in pressure do not influence the remaining concentration. This behaviour reflects that in the later extraction phase, the kinetics of extraction are controlled by the internal diffusion of the solute from solid particles rather than the conditions of the supercritical fluid.

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=0.9\columnwidth]{/Results_sensitivity/CS_P.png}
		\caption{The effect of $P$ change on $C_s$}
		\label{fig:Sensitivty_P_CS}
	\end{figure}

	Figure \ref{fig:Sensitivty_P_CF} shows the sensitivity of the solute concentration in the fluid phase as a response to the pressure change. As explained above, an increase in pressure enhanced the mass transfer, resulting in concentration front moving through the system. Initially, the system response is low due to low response in the solid phase. \todo{TODO}
	Following this, the formation of a trailing front characterized by decreasing sensitivities can be observed. This trailing front can be explained by the decrease in the concentration gradient as the solute is increasingly extracted and the available concentration within the solid phase diminishes. As seen in the figure, sensitivities are declining towards zero. This asymptotic trend suggests the driving force for the mass transfer—the concentration gradient—has been minimized, and changes in pressure no longer significantly impact the solute concentration in the fluid phase.

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=0.9\columnwidth]{/Results_sensitivity/CF_P.png}
		\caption{The effect of $P$ change on $C_f$}
		\label{fig:Sensitivty_P_CF}
	\end{figure}

	Figure \ref{fig:Sensitivty_P_y} illustrates the temporal evolution of the extraction yield's sensitivity to changes in pressure within a supercritical fluid extraction system.

	Initially, the curve displays an almost flat profile, suggesting a latency in the system's response to pressure changes. A minor initial decrease in sensitivity, indicated by a slight dip below zero, may arise due to a decrease in the velocity of the fluid phase. For a constant mass flow rate, an increase in fluid density due to heightened pressure results in a lower fluid velocity, potentially leading to this reduced sensitivity.

	As the process continues, the curve ascends sharply, indicating a positive yield response to pressure. This increase in sensitivity can be associated with the enhanced solubility and mass transfer capabilities imparted by higher pressures, improving the rate at which solute is extracted into the fluid phase. The peak in dy/dP reflects the most responsive period of the system, where pressure changes are highly effective in improving yield.

	Beyond the peak, the sensitivity declines, converging towards zero. This descending curve represents a depletion phase: as the available solute in the solid phase diminishes, the potential for further increasing the yield through pressure alone is reduced. The remaining solute concentration becomes a limiting factor, and the mass transfer driving force, initially augmented by pressure, no longer plays a dominant role.

	\begin{figure}[h!]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=0.9\columnwidth]{/Results_sensitivity/Y_P.png}
		\caption{The effect of $P$ change on $y(t)$}
		\label{fig:Sensitivty_P_y}
	\end{figure}

\end{document}